---
title: "R Notebook"
output: html_notebook
---

#SVM Model

```{r}
kobedata <- read.csv("kobedata.csv", header=TRUE)
head(kobedata)
```

#Move shot_made_flag to the first column, as it is the dependent variable. 
```{r}
shots <- c("shot_made_flag")
kobedata <- kobedata[, c(shots, setdiff(names(kobedata), shots))]
names(kobedata)
```

#Make sure shot_made_flag is a factor in order to run in SVM model. 
```{r}
kobedata$shot_made_flag <- as.factor (kobedata$shot_made_flag)
head(kobedata)
```

#Shot_made_flag contains missing values. The following code finds rows with missing data and removes them. 
```{r}
kobedata <- na.omit(kobedata)
summary(kobedata)
```

#Remove columns we cannot normalize in R. 
```{r}
library(dplyr)
names(kobedata)
```

#Normalize data
```{r}
for (i in 2:dim(kobedata)[2])
{
  kobedata[i] = (kobedata[i]-mean(kobedata[,i]))/sd(kobedata[,i])
}
head(kobedata)
```

#Training and Testing
```{r}
numsamples<-dim(kobedata)[1]
samplingrate=0.8
training <- sample(1:numsamples, samplingrate*numsamples, replace= FALSE)
trainingset <- kobedata[training, ]
testing <-setdiff(1:numsamples, training)
testingset <- kobedata[testing, ]
```

#Load SVM library
```{r}
library(e1071)
```


#Train a support vector machine model using the training set. Apply a linear SVM with an error cost of 20.
```{r}
svmmodel <- svm(shot_made_flag~., data= trainingset, kernel="polynomial", cost=20)
```
#Do predictions on the test set
```{r}
predictions <- predict(svmmodel, testingset, type="class")
predictions
```

#compute misclass rate 
```{r}
sizeoftest = dim(testingset)[1]
error = sum(predictions != testingset$complaints)
missclass = error/sizeoftest
missclass
```
```

